\documentclass{article}
\usepackage{fullpage}
% \usepackage{bibunits}

\setlength\pdfpageheight{\paperheight}
\setlength\pdfpagewidth{\paperwidth}

\makeatletter
\def\@bibunitname{bu-rtm\the\@bibunitauxcnt}
\makeatother

\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phv}

\def\discretionaryslash{\discretionary{/}{}{/}}
{\catcode`\/\active
\gdef\URLprepare{\catcode`\/\active\let/\discretionaryslash
        \def~{\char`\~}}}%
\def\URL{\bgroup\URLprepare\realURL}%
\def\realURL#1{\tt #1\egroup}%

\newcounter{bibctr}

\newcommand{\mkbib}[1]{
    \refstepcounter{bibctr}
    \item[\hskip0pt plus1filll \hbox{[\arabic{bibctr}]}]
  \begin{bibunit}[plain]
      \nocite{#1}
      \putbib[n-str,local,n,n-conf]
    \label{bib:#1}
  \end{bibunit}
}

\newcommand{\refbib}[1]{[\ref{bib:#1}]}

%\renewcommand{\refname}{}

\makeatletter
\renewenvironment{thebibliography}[1]
     {\list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \@openbib@code
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}

\renewcommand{\@biblabel}[1]{}
\countdef\c@nfssec=1
\makeatother

\renewcommand{\thepage}{\Alph{nfssec}-\arabic{page}}
\def\up#1{\raise.16ex\hbox{#1}}



\begin{document}
\count1=6 % This sets page numbers to be F-n
\count0=3 % This first page to be F-n
\section*{Biographical Sketch: Robert Morris}

\subsection*{(a) Professional Preparation}

\begin{tabbing}
Massachusetts Inst.\ of Technology \quad \=
    Elec. Eng. and Comp. Sci. \quad \=
    Postdoctoral Scholar \quad \=
    \kill

\textbf{Institution} \> \textbf{Major} \> \textbf{Degree} \> \textbf{Date} \\
Harvard University \> Computer Science \> B.A. \> 1/1988 \\
Harvard University \> Computer Science \> Ph.D. \> 1/1999 \\
\end{tabbing}

\subsection*{(b) Appointments}

\begin{tabbing}
Massachusetts Inst.\ of Technology\quad \=
    Elec. Eng. and Comp. Sci. \quad \=
    \kill
\textbf{Organization}\> \textbf{Position}\> \textbf{Dates}\\

Massachusetts Inst.\ of Technology \> Professor \> 2010-- \\
Massachusetts Inst.\ of Technology \> Associate Professor \> 2003--2010 \\
Massachusetts Inst.\ of Technology \> Assistant Professor \> 2000--2003 \\

\end{tabbing}

%\paragraph{\textbf{Honors:}}
%Alfred P. Sloan Research Fellow (2003), NSF CAREER Award (2003), ITT
%Career Development Chair (2003), award papers at USENIX
%1994, SOSP 1999, SIGCOMM 2004, SOSP 2005, MobiCom 2005, and SIGCOMM 2005.

\subsection*{(c) Publications}

\subsubsection*{Publications Relevant to Proposal}
\begin{itemize}

\item Yandong Mao, Eddie Kohler, and Robert Morris. Cache Craftiness
  for Fast Multicore Key-Value Storage. In the Proceedings of the ACM
  EuroSys Conference (EuroSys 2012), Bern, Switzerland, April 2012.

\item Silas Boyd-Wickizer, Austin T. Clements, Yandong Mao, Aleksey
  Pesterev, M. Frans Kaashoek, Robert Morris, and Nickolai Zeldovich.
  An Analysis of Linux Scalability to Many Cores. In the Proceedings
  of the 9th USENIX Symposium on Operating Systems Design and
  Implementation (OSDI '10), Vancouver, Canada, October 2010.

\item Aleksey Pesterev, Nickolai Zeldovich, and Robert T. Morris.
  Locating Cache Performance Bottlenecks Using Data Profiling. In the
  Proceedings of the ACM EuroSys Conference (EuroSys 2010), Paris,
  France, April 2010.

\item Silas Boyd-Wickizer, Haibo Chen, Rong Chen, Yandong Mao, Frans
  Kaashoek, Robert Morris, Aleksey Pesterev, Lex Stein, Ming Wu,
  Yuehua Dai, Yang Zhang, and Zheng Zhang. Corey: An Operating System
  for Many Cores. In the Proceedings of the 8th USENIX Symposium on
  Operating Systems Design and Implementation (OSDI 2008), San Diego,
  California, December 2008, pages 43-57.

\item Benjie Chen and Robert Morris. Flexible control of parallelism
  in a multiprocessor PC router. In Proceedings of the 2001 USENIX
  Annual Technical Conference (USENIX 2001), Boston, Massachusetts,
  June 2001.

\end{itemize}

\subsubsection*{Selected Other Publications}
\begin{itemize}

\item Silas Boyd-Wickizer, M. Frans Kaashoek, Robert Morris, and
  Nickolai Zeldovich. Non-scalable locks are dangerous. In the
  Proceedings of the Linux Symposium, Ottawa, Canada, July 2012.

\item Aleksey Pesterev, Jacob Strauss, Nickolai Zeldovich, and Robert
  T. Morris. Improving Network Connection Locality on Multicore
  Systems. In the Proceedings of the ACM EuroSys Conference (EuroSys
  2012), Bern, Switzerland, April 2012.

\item Silas Boyd-Wickizer, Robert Morris, and M. Frans Kaashoek.
  Reinventing scheduling for multicore systems. In Proceedings of the
  12th Workshop on Hot Topics in Operating Systems (HotOS-XII), Monte
  Verita, Switzerland, May 2009.

\item Nickolai Zeldovich, Alexander Yip, Frank Dabek, Robert T.
  Morris, David Mazieres, and M. Frans Kaashoek. Multiprocessor
  support for event-driven programs. In Proceedings of the USENIX 2003
  Annual Technical Conference, San Antonio, TX, June 2003.

\end{itemize}

\subsection*{(d) Synergistic Activities}

\begin{itemize}

\item Developing and teaching a graduate course on building robust and
high-performance distributed systems.
\item Program committee member for SIGCOMM 2002, OSDI 2002, Mobicom
  2002, SIGCOMM 2003, SIGCOMM 2004, FAST 2005, SIGCOMM 2006, SIGCOMM
  2008, NSDI 2008, FAST 2008, SIGCOMM 2008, SIGCOMM 2010, Mobicom 2010,
  SOSP 2011, OSDI 2012. Co-chair of Program Committee for MobiSys
  2003, NSDI 2004, and HotNets 2010.
\end{itemize}

\subsection*{(e) Collaborators \& Other Affiliations}
\noindent\textbf{Non-MIT Collaborators:}
    Edward Kohler (Harvard).

\noindent\textbf{Graduate Advisor:}
    H. T. Kung (Harvard).

\noindent\textbf{Student Advisees:}
    Edward Kohler (Harvard),
    Benjie Chen,
    Douglas De Couto,
    Jinyang Li (NYU),
    Frank Dabek (Google),
    Emil Sit (VMWare),
    Max Krohn (OK Cupid),
    Athicha Muthitacharoen,
    Jeremy Stribling (Nicira Networks),
    Alex Yip (Google),
    Jayashree Subramanian (Bell Labs India),
    Silas Boyd-Wickizer,
    Alex Pesterev,
    Neha Narula,
    Yandong Mao,
    Cody Cutler.
    Ph.D. Total: 11~completed, 5~current.

\end{document}
