\documentclass{article}
\usepackage{fullpage}
\usepackage{bibunits}

\setlength\pdfpageheight{\paperheight}
\setlength\pdfpagewidth{\paperwidth}

\makeatletter
\def\@bibunitname{bu-nickolai\the\@bibunitauxcnt}
\makeatother

\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phv}

\def\discretionaryslash{\discretionary{/}{}{/}}
{\catcode`\/\active
\gdef\URLprepare{\catcode`\/\active\let/\discretionaryslash
        \def~{\char`\~}}}%
\def\URL{\bgroup\URLprepare\realURL}%
\def\realURL#1{\tt #1\egroup}%

\newcounter{bibctr}

\newcommand{\mkbib}[1]{
    \refstepcounter{bibctr}
    \item[\hskip0pt plus1filll \hbox{[\arabic{bibctr}]}]
  \begin{bibunit}[plain]
      \nocite{#1}
      \putbib[n-str,n,n-conf]
    \label{bib:#1}
  \end{bibunit}
}

\newcommand{\refbib}[1]{[\ref{bib:#1}]}

%\renewcommand{\refname}{}

\makeatletter
\renewenvironment{thebibliography}[1]
     {\list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \@openbib@code
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}

\renewcommand{\@biblabel}[1]{}
\countdef\c@nfssec=1
\makeatother

\renewcommand{\thepage}{\Alph{nfssec}-\arabic{page}}
\def\up#1{\raise.16ex\hbox{#1}}



\begin{document}
\count1=6 % This sets page numbers to be F-n
\count0=1 % This first page to be F-n
\section*{Biographical Sketch: Nickolai Zeldovich}

\subsection*{(a) Professional Preparation}

\begin{tabbing}
Massachusetts Inst.\ of Technology \quad \=
    Elec. Eng. and Comp. Sci. \quad \=
    Postdoctoral Scholar \quad \=
    \kill

\textbf{Institution} \> \textbf{Major} \> \textbf{Degree} \> \textbf{Date} \\
Massachusetts Inst.\ of Technology \> Elec. Eng. and Comp. Sci. \> S.B. \> 6/2002 \\
Massachusetts Inst.\ of Technology \> Elec. Eng. and Comp. Sci. \> M.Eng. \> 6/2002 \\
Stanford University \> Computer Science \> Ph.D. \> 1/2008 \\
Stanford University \> Computer Science \> Postdoctoral Scholar \> 1/2008--7/2008 \\
\end{tabbing}

\subsection*{(b) Appointments}

\begin{tabbing}
Massachusetts Inst.\ of Technology\quad \=
    Elec. Eng. and Comp. Sci. \quad \=
    \kill
\textbf{Organization}\> \textbf{Position}\> \textbf{Dates}\\

Massachusetts Inst.\ of Technology \> Associate Professor \> 7/2012-- \\
Massachusetts Inst.\ of Technology \> Assistant Professor \> 7/2008--6/2012 \\
moka5 \> Co-founder \> 5/2005--9/2005 \\
Kaveri Networks \> Software Developer \> 6/2000--8/2000 \\
Naval Research Laboratory \> Summer Intern \> 6/1999--8/1999 \\

\end{tabbing}

%\textbf{Honors:} Douglas T. Ross Career Development Chair (2009--2012).

\subsection*{(c) Publications}

\subsubsection*{Publications Relevant to Proposal}
\begin{itemize}
    \mkbib{clements:bonsai}
    \mkbib{boyd-wickizer:scaling}
    \mkbib{boyd-wickizer:locks}
    \mkbib{pesterev:multiaccept}
    \mkbib{pesterev:dprof}
\end{itemize}

\subsubsection*{Selected Other Publications}
\begin{itemize}
    \mkbib{zeldovich:histar}
    \mkbib{popa:cryptdb}
    \mkbib{kim:retro}
    \mkbib{chandra:collective}
    \mkbib{yip:resin}
\end{itemize}

\subsection*{(d) Synergistic Activities}

\begin{itemize}
\item Program committee member for SOSP 2009, Usenix 2010, OSDI 2010, SOSP 2011, NSDI 2012, OSDI 2012, ASPLOS 2013, SOSP 2013.
\item Teaching courses on computer systems (6.033), operating systems (6.828),
      and computer systems security (6.858) at MIT.
\end{itemize}

\subsection*{(e) Collaborators \& Other Affiliations}
\paragraph{Non-MIT Collaborators:}
    Michael Dalton (Stanford),
    Eddie Kohler (Harvard),
    Christos Kozyrakis (Stanford),
    Philip Levis (Stanford),
    David Mazi\`eres (Stanford),
    Nick McKeown (Stanford),
    Arjun Roy (Stanford),
    Stephen Rumble (Stanford),
    Ryan Stutsman (Stanford).

\paragraph{Graduate Advisors:}
    Monica Lam (Stanford),
    David Mazi\`eres (Stanford),
    Robert Morris (MIT).

\paragraph{Postdoctoral Sponsor:}
    David Mazi\`eres (Stanford).

\paragraph{Student Advisees:}
    Ramesh Chandra (MIT),
    Haogang Chen (MIT),
    Austin Clements (MIT),
    Priya Gupta (Google),
    Taesoo Kim (MIT),
    Raluca Ada Popa (MIT),
    Meelap Shah (MIT),
    Emily Stark (MIT),
    Frank Wang (MIT),
    Hsin-Jung Yang (MIT),
    Alex Yip (Nicira).

    Student total: 3~completed, 8~current.

\paragraph{Postdocs:}
    Jad Naous (Nerati).

    Postdoc total: 1~completed, 0~current.

\end{document}
