\documentclass[11pt]{article}
\usepackage{url,mathptmx,times,verbatim}
\usepackage[hmargin=1in,vmargin=1in,letterpaper]{geometry}
\frenchspacing

\renewcommand{\thesection}{\Alph{section}}
\renewcommand{\thesubsection}{\Alph{section}.\alph{subsection}}

%\input{fonts.sty}

\begin{document}

\setcounter{section}{4}

\section*{Biographical Sketch: Eddie Kohler}

\subsection*{(a) Professional Preparation}

\begin{flushleft}
\begin{tabular}{@{}lll@{}}
Undergraduate & Massachusetts Institute of Technology \\
	& Mathematics with Computer Science	& S.B., 1995 \\
	& Music					& S.B., 1995 \\
Graduate & Massachusetts Institute of Technology \\
	& Electrical Engineering and Computer Science & S.M., 1998;
						Ph.D., 2001 \\
Postdoctoral & International Computer Science Institute, Berkeley \\
	& Computer Science			& 2001--2003
\end{tabular}
\end{flushleft}


\subsection*{(b) Appointments}

\begin{flushleft}
\begin{tabular}{@{}r@{}ll@{}}
7&/11--present & Associate Professor, Harvard University \\
10&/09--6/10 & Visiting faculty, Intel Labs Berkeley \\
4&/08--7/10 & Senior member of technical staff, Meraki, Inc. \\
1&/04--7/11 & Assistant and Associate Professor, University of California, Los Angeles
\\
2&/03--9/03 & Member of research staff, ICSI Center for Internet
Research \\
10&/00--2/09 & Chief Scientist, Mazu Networks, Inc. \\
\end{tabular}
\end{flushleft}

\subsection*{(c) Publications}

\subsubsection*{Publications Relevant to Proposal}

\begin{itemize}
\itemsep=0pt
\parsep=0pt
\leftmargin=0pt
\labelwidth=0pt

\item Yandong Mao, Eddie Kohler, and Robert Morris.  ``Cache craftiness for
  multicore key-value storage.'' In \emph{Proc. 7th ACM European Conference
    on Computer Systems (EuroSys '12)}, April 2012.

\item Mike Mammarella, Shant Hovsepian, and Eddie Kohler. ``Modular data
  storage with Anvil.'' In \emph{Proc. 22nd ACM Symposium on Operating
  Systems Principles (SOSP '09)}, pages 147--160, October 2009.

\item Steve VanDeBogart, Christopher Frost, and Eddie Kohler. ``Reducing
  seek overhead with application-directed prefetching.'' In \emph{Proc. 2009
  USENIX Annual Technical Conference}, June 2009.

\item Christopher Frost, Mike Mammarella, Eddie Kohler, Andrew de los
  Reyes, Shant Hovsepian, Andrew Matsuoka, and Lei Zhang. ``Generalized
  file system dependencies.'' In \emph{Proc. 21st ACM Symposium on
  Operating Systems Principles (SOSP '07)}, pages 307--320, October 2007.

\item Maxwell Krohn, Eddie Kohler, and M. Frans Kaashoek. ``Events can make
  sense.'' In \emph{Proc. 2007 USENIX Annual Technical Conference}, pages
  87--100, June 2007.

\end{itemize}

\subsubsection*{Selected Other Publications}

\begin{itemize}
\itemsep=0pt
\parsep=0pt
\leftmargin=0pt
\labelwidth=0pt

\item Steve VanDeBogart, Petros Efstathopoulos, Eddie Kohler, Maxwell
  Krohn, Cliff Frey, David Ziegler, Frans Kaashoek, Robert Morris, and
  David Mazi\`eres. ``Labels and event processes in the Asbestos operating
  system.'' \emph{ACM Transactions on Computer Systems}
  \textbf{25}(4):11:1--43, November 2007.

\item Nithya Ramanathan, Thomas Schoellhammer, Eddie Kohler, Kamin
  Whitehouse, Thomas Harmon, and Deborah Estrin. ``Suelo: Human-assisted
  sensing for exploratory soil monitoring studies.'' In \emph{Proc. 7th
  International Conference on Embedded Networked Sensor Systems (ACM SenSys
  '09)}, pages 197--210, November 2009.

\item Mark Handley, Eddie Kohler, Atanu Ghosh, Orion Hodson, and Pavlin
  Radoslavov. ``Designing extensible IP router software.'' In
  \emph{Proc.\ 2nd Symposium on Networked Systems Design \& Implementation
  (NSDI '05)}, May 2005.

\item Benjamin Greenstein, Eddie Kohler, and Deborah Estrin.  ``A Sensor
  Network Application Construction Kit (SNACK).''  In \emph{Proc. 2nd ACM
  Conference on Embedded Network Sensor Systems (SenSys~'04)}, November
  2004.

\item Eddie Kohler, Robert Morris, Benjie Chen, John Jannotti, and M. Frans
  Kaashoek. ``The Click modular router.'' \emph{ACM Transactions on
  Computer Systems} \textbf{18}(3), August 2000, pages 263--297.

\end{itemize}


\subsection*{(d) Synergistic Activities}
%
\textbf{Program committee member:} ACM SOSP 2013, Usenix 2012 Annual
Technical Conference (ATC), Usenix NSDI 2012, ACM SOSP 2011, Usenix 2011
ATC, Usenix NSDI 2010, ACM HotNets 2010, ACM SIGCOMM 2009, Usenix NSDI
2008, ASPLOS 2008, Usenix 2007 ATC, IPSN 2007, EmNets 2006 (partial
list). Program committee co-chair: ACM HotNets 2006.

\textbf{Awards:} US Presidential Early Career Award for Scientists and
Engineers (PECASE, nominated by NSF), 2007; Sloan Research Fellowship,
2007; Microsoft Research New Faculty Fellowship, 2006; UCLA Henry Samueli
Fellowship, 2004--2007.

\textbf{Software:} Maintains and supports many widely-used open-source
software distributions for networking and systems research. The Click
modular router (\url{http://read.cs.ucla.edu/click/}) is very widely
used in research and industry; tens of thousands of deployed networking
devices use Click technology. HotCRP
(\url{http://read.seas.harvard.edu/~kohler/hotcrp/}) supports academic
conference review management and is used by the USENIX Association for
all its conferences. Ipsumdump, lcdf-typetools, gifsicle, and other
packages are also widely used.

\textbf{Classes:} Developing and teaching undergraduate and graduate
courses on systems, including graduate courses on readable operating
system kernels, distributed systems infrastructure, and operating
systems design.

\subsection*{(e) Collaborators \& Other Affiliations} 

\noindent\textbf{Collaborators:} 
Katerina Argyraki (EPFL),
%Milan Bradonjic (UCLA),
%Byung-Gon Chun (Intel Labs Berkeley),
%Kevin Fall (Intel Labs Berkeley),
%Allan Knies (Intel),
Mihai Dobrescu (EPFL),
Norbert Egi,
Deborah Estrin (UCLA),
Kevin Fall (Qualcomm),
%Sally Floyd (ICIR),
%Atanu Ghosh (ICIR),
Ramesh Govindan (USC),
Benjamin Greenstein (Google),
%Dirk Grunwald (University of Colorado),
%Mark Handley (University College London),
%Thomas Harmon (UC Merced),
%Orion Hodson (Microsoft Research),
Gianluca Iannaconne (Red Bow Labs),
%Ranjit Jhala (UCSD),
M.~Frans Kaashoek (MIT),
%Dina Katabi (MIT),
Jinyang Li (NYU),
Barbara Liskov (MIT),
%Rupak Majumdar (UCLA),
Maziar Manesh (Intel),
Todd Millstein (UCLA),
Robert Morris (MIT),
Rafail Ostrovsky (UCLA),
%Jitendra Padhye (Microsoft Research),
%Jens Palsberg (UCLA),
%Vern Paxson (ICIR),
%Greg Pottie (UCLA),
%Massimiliano Poletto (Google),
%Pavlin Radoslavov (ICIR),
Sylvia Ratnasamy (UC Berkeley)
%Niraj Shah (UC Berkeley),
%Scott Shenker (ICIR/UC Berkeley),
%Mani Srivastava (UCLA),
%Kamin Whitehouse (University of Virginia),
%Lixia Zhang (UCLA)

\textbf{Graduate and Postdoctoral Advisors:}
M.~Frans Kaashoek (MIT),
Robert Morris (MIT).

\textbf{Student Advisees:}
Thesis advisor or co-advisor of:
Russ Cox (co-advisor; Google),
Petros Efstathopoulos (Symantec Research),
Christopher Frost (Google),
Benjamin Greenstein (co-advisor; Google),
Maxwell Krohn (co-advisor; OK Cupid),
Nikitas Liogkas,
Michael Mammarella (Google),
Nithya Ramanathan (co-advisor; NexLeaf/UCLA Center for Embedded Network Sensing),
Steven VanDeBogart (Google),
Bryan Kate,
Michael Kester.
Ph.D. Total: 9~completed, 2~current.

\end{document}

