\documentclass{article}
\usepackage{url}
\usepackage{fullpage}
\usepackage{bibunits}

\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phv}

\setlength\pdfpageheight{\paperheight}
\setlength\pdfpagewidth{\paperwidth}

\def\discretionaryslash{\discretionary{/}{}{/}}
{\catcode`\/\active
\gdef\URLprepare{\catcode`\/\active\let/\discretionaryslash
        \def~{\char`\~}}}%
\def\URL{\bgroup\URLprepare\realURL}%
\def\realURL#1{\tt #1\egroup}%

\renewcommand{\thepage}{F-\arabic{page}}

\begin{document}
\count1=6 % This sets page numbers to be F-n
\count0=3 % This first page to be F-n
\section*{Biographical Sketch: M. Frans Kaashoek}

\subsection*{(a) Professional Preparation}

\begin{tabular}{p{2in}p{1.5in}p{1in}p{1in}}
{\bf Institution} & {\bf Major} & {\bf Degree} & {\bf Date}\\
Vrije Universiteit (Amsterdam) & Computer Science & Ph.D. & 1992 \\
\end{tabular}

\subsection*{(b) Appointments}

\begin{tabular}{p{1.25in}p{2in}p{2in}}
{\bf Organization} & {\bf Position} & {\bf Dates}\\
MIT EECS & Full Professor & 2001 -\\
MIT EECS & Associate Professor & 1996 - 2001 \\
MIT EECS & Assistant Professor & 1993 - 1996 \\
\end{tabular}

\vspace{0.1in}

%\textbf{Honors:} NSF national young investigator award (1994);
%Jamieson Career Development chair (1995);
%USENIX technical conference best paper award (1997);
%SOSP conference best paper award (1999);
%OSDI conference best paper award (2000);
%MIT IEEE best undergraduate advisor award (2000);
%ACM Mark Weiser award (2001);
%ACM Fellow (2004);
%SOSP conference best paper award (2005);
%National Academy of Engineering (2006).

\subsection*{(c) Publications}

\subsubsection*{Publications Relevant to Proposal}

\begin{enumerate}

\item Silas Boyd-Wickizer and Haibo Chen and Rong Chen and Yandong Mao and Frans
  Kaashoek and Robert Morris and Aleksey Pesterev and Lex Stein and Ming Wu and
  Yuehua Dai and Yang Zhang and Zheng Zhang.  Corey: An Operating System for
  Many Cores, Proceedings of the 8th {USENIX} {S}ymposium on {O}perating
  {S}ystems {D}esign and {I}mplementation ({OSDI} '08), 2008.

\item Silas Boyd-Wickizer, Austin T. Clements, Yandong Mao, Aleksey Pesterev,
  M. Frans Kaashoek, Robert Morris, and Nickolai Zeldovich. An Analysis of Linux
  Scalability to Many Cores. In the Proceedings of the 9th USENIX Symposium on
  Operating Systems Design and Implementation (OSDI '10), Vancouver, Canada,
  October 2010.

\item Silas Boyd-Wickizer, M. Frans Kaashoek, Robert Morris, and Nickolai
  Zeldovich. A Software Approach to Unifying Multicore Caches. MIT Computer
  Science and Artificial Intelligence Laboratory technical report
  MIT-CSAIL-TR-2011-032, June 2011.

\item Austin T. Clements, M. Frans Kaashoek, Nickolai Zeldovich.  Scalable
  Address Spaces Using RCU Balanced Trees. In the Proceedings of the 17th
  International Conference on Architectural Support for Programming Languages
  and Operating Systems (ASPLOS), London, UK, March 2012.

\item Silas Boyd-Wickizer, M. Frans Kaashoek, Robert Morris, and Nickolai
  Zeldovich. Non-scalable locks are dangerous. In the Proceedings of the Linux
  Symposium, Ottawa, Canada, July 2012.

\end{enumerate}

\subsubsection*{Selected Other Publications}

\begin{enumerate}{
  \setlength{\topsep}{-10pt}
  \setlength{\itemsep}{0pt}
  \setlength{\parskip}{0pt}}

\item Information Flow Control for Standard OS Abstractions. Maxwell
  Krohn and Alexander Yip and Micah Brodsky and Natan Cliffer and
  M. Frans Kaashoek and Eddie Kohler and Robert Morris, Proceedings
  of the 21st Symposium on Operating Systems Principles, 2007.

\item Ion Stoica, Robert Morris, David Karger, M. Frans Kaashoek, and
  Hari Balakrishnan, Chord: A Scalable Peer-to-peer Lookup Service for
  Internet Applications, ACM SIGCOMM 2001, San Deigo, CA, August 2001.

\item David Andersen, Hari Balakrishnan, M. Frans Kaashoek, and Robert
  Morris, Resilient Overlay Networks, ACM SOSP 2001, Banff, October
  2001.

\item Eddie Kohler, Robert Morris, Benjie Chen, John Jannotti, and
  M. Frans Kaashoek, The Click Modular Router, ACM Transactions on
  Computer Systems, August 2000, 18(4), pages 263-297.

\item D. R. Engler, M. F. Kaashoek, and J. O'Toole Jr. Exokernel: an
  operating system architecture for application-specific resource
  management. In Proceedings of the Fifteenth ACM Symposium on
  Operating Systems Principles, December 1995.

\end{enumerate}

\subsection*{(d) Synergistic Activities}

\begin{enumerate}

\item MIT course on operating systems (6.828).

\item Contributor to various public-domain systems that are deployed.

\item Chair and PC member of various top conferences in computer
systems and networking.

\end{enumerate}

\subsection*{(e) Collaborators \& Other Affiliations}
\noindent\textbf{Non-MIT Collaborators:}  David Andersen (CMU),
Wenguang Chen (Tsinghua University), Peter Druschel (MPI-SWS), Joe
Hellerstein (Berkeley), Eddie Kohler (UCLA), John Kubiatowicz, David
Mazieres (NYU/Stanford), Scott Shenker (ICSI/Berkeley), Ion Stoica
(Berkeley), Zheng Zhang (MSR Asia).

\vspace{0.1in}

\noindent\textbf{Graduate Advisors:} Andy Tanenbaum (Vrije
Universiteit, NL).

\vspace{0.1in}

\noindent\textbf{Thesis (co)advisor:}
Jacob Strauss (QRC), Chris Lesniewski-Laas (Google), Jeremy Stribling (Nicira),
Max Krohn (OkCupid), Russ Cox (Google), Bryan Ford (Yale), Emil Sit (VMware),
Frank Dabek (Google), Kevin Fu (U. Amherst), Michael Kaminsky (Intel Research),
Alex Snoeren (UCSD), John Jannotti (Brown), David Mazieres (Stanford), Eddie
Kohler (UCLA/Meraki), Massimiliano Polletto (Meraki), Dawson Engler (Stanford),
Deborah Wallach (Google), Anthony Joseph (UCB), Ken Mackenzie (Unknown), Wilson
Hsieh (Google), Kirk Johnson (Google), Sanjay Ghemawat (Google).

Ph.D. Total: 22~completed, 4~current.

\vspace{0.1in}

\noindent\textbf{Postgraduate-scholar Sponsor:} Greg Ganger (CMU), Max
Poletto (Google), John Jannotti (Brown), Robert Morris (MIT), Sean
Rhea (Meraki).

Postdoc Total: 5~completed.


\end{document}
