\section{Expected contributions}

The expected contributions of this thesis are as follows.

\begin{itemize}
\item This thesis will provide a formal definition and a proof of the
  basic commutativity rule, as well as variants of the rule tailored
  to common interface properties found in practice.  This will include
  a formalized model of scalability designed to capture hardware
  capabilities.  Our hope is that this rule and its variants will
  enable software practitioners to apply modular scalability reasoning
  to the design of software interfaces, the realization of scalable
  implementations of these interfaces, and the use of such interfaces.

\item This thesis will demonstrate the application of the
  commutativity rule to interface design by exploring its implications
  for the POSIX system call API and how this API can be changed to
  improve implementation scalability.  We will demonstrate through
  analysis and benchmarks that the resulting interface modifications
  enable highly scalable implementations of these interfaces.

\item This thesis will develop a new operating system, \sys, to
  demonstrate the application of the commutativity rule to a large
  software system.  This will include the development of new software
  techniques as well as new applications of existing software
  techniques.

\item This thesis will develop techniques and a corresponding tool for
  verifying that an implementation of an interface satisfies the
  commutativity rule.
\end{itemize}
