% If any PI or co-PI identified on the project has received NSF funding in
% the past five years, information on the award(s) is required. Each PI
% and co-PI who has received more than one award(excluding amendments)
% must report on the award most closely related to the proposal. The
% following information must be provided:
% 
% (a) the NSF award number, amount and period of support;
% 
% (b) the title of the project;
% 
% (c) a summary of the results of the completed work, including, for a
% research project, any contribution to the development of human resources
% in science and engineering;
% 
% (d) publications resulting from the NSF award;
% 
% (e) a brief description of available data, samples, physical collections
% and other related research products not described elsewhere; and
% 
% (f) if the proposal is for renewed support, a description of the relation
% of the completed work to the proposed work.

\section{Results From Prior Support for PIs}
\label{sec:prior}

\textit{NSF CSR, Medium: Collaborative Research: Programming parallel in-memory
  data-center applications with Piccolo} (CNS 1065114, 4/11--5/14,
\$447,852).
[Kaashoek]
This project investigates a new parallel programming model for data
centers. MIT's part is to investigates how to implement this model on multicore
computers.

\textit{NSF SHF, Medium: Intelligent and efficient data movement for
  multicore systems} (SHF 0964106, 6/10--6/13, \$1,080,000).
[Kaashoek, Morris, Zeldovich]
This
project explores a novel hardware/software interface that helps
applications control placement and movement of data with respect to
on-chip caches and networks in multicore chips.  We have explored
software techniques to mitigate the cost of cache line movement in
multi-core processors~\cite{pesterev:dprof, pesterev:multiaccept,
metreveli:cphash-poster}.

\textit{NSF CSR: Small: CoreTime: Dynamic Computation Migration for Multicore
  System Software} (8/09--8/12, \$500,000).
[Kaashoek, Morris]
This proposal investigates data
structure design for parallel systems.  One result of this work is a
virtual-memory design that allows threads in the same address space to perform
concurrent read operations on the in-kernel VM data
structures~\cite{clements:bonsai}.  We would like to generalize those ideas to
support concurrent write-write operations (such as \code{unmap} and \code{map}
that commute because they apply to different parts of a shared address space).
Papers and software are available at \URL{http://pdos.csail.mit.edu/multicore/}.

\textit{NSF CSR-PSCE, SM: An operating system for multicore processors} (CNS
0834415, 8/08--8/11, \$450,000).
[Kaashoek, Morris]
Corey is a new operating system for many core
processors, which supports new operating system abstractions that help
applications avoid scalability bottlenecks in the operating system~\cite{boyd-wickizer:corey}.
The benchmarks we developed for Corey have been used by others, including our
study of Linux scalability~\cite{boyd-wickizer:scaling}.  As part of the Corey and
CoreTime proposal, we developed xv6~\cite{xv6}, which has been adopted by
several schools for educational and research purposes.

\textit{NSF CNS, CAREER: Fine-Grained Operating System Components for
  Storage} (CNS~0546892, 2/06--1/11, \$400,000). [Kohler] This
proposal laid out a research program for representing file systems as
collections of components, organized around a generalized notion of
file system \emph{dependency}. The file system component
infrastructure and other related projects, including a prototype
system for structured storage (Anvil) and operating system extensions
for prefetching (libprefetch), required extensive modeling of
concurrent applications to improve performance~\cite{mammarella09modular,vandebogart09reducing,liogkas08exploring,chun08netcomplex,frost07generalized,delosreyes05kudos}.

% LocalWords:  MSRA Kaashoek DHTs DHT IPTPS com Nokia JOS CourseWare UIA NeTS
% LocalWords:  SourceForge serverless Pastwatch CVS

% LocalWords:  MSRA Kaashoek DHTs DHT IPTPS com Nokia JOS CourseWare UIA NeTS
% LocalWords:  SourceForge serverless Pastwatch CVS
