\chapter*{Acknowledgments}

This work would not have been possible without the combined efforts
and determination of my advisors Frans Kaashoek, Nickolai Zeldovich,
Robert Morris, and Eddie Kohler.
%
Frans' unwavering focus on The Technical Nugget led to our simplest
and most powerful ideas.
%
Nickolai's raw energy and boundless breadth are reflected in every
corner of this dissertation.
%
Robert's keen vision and diligent skepticism pushed the frontiers of
what we thought possible.
%
And Eddie's inexhaustible enthusiasm and remarkable intuition reliably
yielded solutions to seemingly impossible problems.
%
I am honored to have had them as my mentors. % for nearly ten years now.

This research was greatly improved by the valuable feedback of Silas
Boyd-Wickizer, Yandong Mao, Xi Wang, Marc Shapiro, Rachid Guerraoui,
Butler Lampson, Paul McKenney, and the twenty three anonymous
reviewers who pored over the many drafts of this work.
% SOSP 5, OSDI 9, RadixVM 4, Bonsai 5

Silas and Nickolai contributed substantially to the implementation of
\sys and \sys would not have been possible without xv6, which was
written by Frans, Robert, and Russ Cox.

I thank my parents, Hank and Robin, for their encouragement, guidance,
and support. %, and for setting me on my path.
%
Finally, I thank my wife, Emily, for her love and understanding, and
for being my beacon through it all.

\asterism

This research was supported by NSF awards SHF-964106 and CNS-1301934,
by grants from Quanta Computer and Google, and by a VMware graduate
fellowship.

\asterism

This dissertation incorporates and extends work previously published
in the following papers:

\begin{quote}
  Austin~T. Clements, M.~Frans Kaashoek, Nickolai Zeldovich,
  Robert~T. Morris, and Eddie Kohler.
  \newblock The scalable commutativity rule: Designing scalable
  software for multicore processors.
  \newblock In \emph{Proceedings of the 24th ACM Symposium on
    Operating Systems Principles (\mbox{SOSP})}, Farmington,
  Pennsylvania, November 2013.

  Austin~T. Clements, M.~Frans Kaashoek, and Nickolai Zeldovich.
  \newblock {RadixVM}: Scalable address spaces for multithreaded
  applications.
  \newblock In \emph{Proceedings of the ACM EuroSys Conference},
  Prague, Czech Republic, April 2013.

  Austin~T. Clements, M.~Frans Kaashoek, and Nickolai Zeldovich.
  \newblock Concurrent address spaces using {RCU} balanced trees.
  \newblock In \emph{Proceedings of the 17th International Conference
    on Architectural Support for Programming Languages and Operating
    Systems (\mbox{ASPLOS})}, London, UK, March 2012.
\end{quote}
